Towards a Mechanized Proof of Noninterference

### 2013-12-19 paper/paper.tex
earlyoften

In this paper we will present our simplified language, \langName,
along with its semantics and type system (Section~\ref{sec:lang}).  We will then
present a natural language proof of noninterference for the
language and discuss our experiences with mechanizing this proof in the Coq
proof assistant (Sections~\ref{sec:proof} and \ref{sec:mechanizing}).  Finally,
we will discuss related work, next steps, and conclude
(Sections~\ref{sec:related}, \ref{sec:future}, and \ref{sec:conclusion}).

@@

Our toy language for proving noninterference, \langName, is a simplified version
of \origLang from Simonet and Pottier's paper~\cite{InfoFlowML}.  Since we had a
limited amount of time to work on this project, we restricted ourselves to a subset
of the language that drops the following features:

\begin{itemize}
  \item References and associated memory operations.
  \item Value Tuples.
  \item Exceptions.
  \item Sequencing operations.
\end{itemize}

While the language becomes much more burdensome for the programmer, we
feel that these features do not fundamentally change the problem of proving
noninterference.  Mechanizing a proof of noninterference for a
language closer to \origLang would require simply adding mechanized proofs
for the additional cases where appropriate.

@@

A reader familiar with the type system for \origLang will notice that our type
system does not include labels above the arrow in the function type, which would
normally give security bounds on side effects of the function, nor does it
include effect labels for exceptions.  These are dropped as a result of the
simplifications we made in defining \langName, since there are no side effects
or exceptions in the language.

>>>>

In this paper we will present our simplified language, \langName,
along with its semantics and type system (Section~\ref{sec:lang}).  We will then
present a natural language, human readable, proof of noninterference for the
language and discuss our experiences with mechanizing this proof in the Coq
proof assistant (Sections~\ref{sec:proof} and \ref{sec:mechanizing}).  Finally,
we will discuss related work, next steps, and conclude
(Sections~\ref{sec:related}, \ref{sec:future}, and \ref{sec:conclusion}).

@@

Our toy language for proving noninterference, \langName, is a simplified version
of \origLang from Simonet and Pottier's paper~\cite{InfoFlowML}.  Since we had a
limited amount of time to work on this project, we limited ourselves to a subset
of the language that drops the following features:

\begin{itemize}
  \item References and associated memory operations.
  \item Paired values.
  \item Exceptions.
  \item Sequencing operations.
\end{itemize}

While the language becomes much much less interesting for actual programming, we
feel that these features do not fundamentally change the problem of proving
noninterference.  In order to mechanize a proof of noninterference for a
language closer to \origLang, it would require simply adding mechanized proofs
for the additional cases where appropriate.

@@

A reader familiar with the type system for \origLang will notice that our type
system does not include types above the arrow in the function type, for giving
security bounds on side effects of the function, and the type does not include
effect labels for exceptions.  These are dropped as a result of the
simplifications we made in defining \langName, since there are no side effects
or exceptions in the language.
